---
title: "BDA- Assignment 1"
author: "Anonymous"
date: "08/03/2021"
output:
    pdf_document:
    toc: yes
    toc_depth: 1
---

```{r setup, include=FALSE}
# This chunk sets echo = TRUE as default, that is print all code.
# knitr::opts_chunk$set can be used to set other notebook generation options, too.
# include=FALSE inside curly brackets makes this block not be included in the pdf.
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# To install aaltobda, see the General information in the assignment.
library(aaltobda)
```

# Exercise 1)

* probability: is a measure of uncertainty; the chance of an event happening.
* probability mass: is a point in the distribution of a pmf.
* probability density: is a point in the distribution of a pdf.
* probability mass function: is a function that takes as an input a value in the range of a discrete random variable, and outputs a probability for that event.
* probability density function: is a function that takes as an input a value in the range of a continuous random variable, and outputs a probability for that event.
* probability distribution: is a function, a pmf or a pdf.
* discrete probability distribution: a probability distribution that takes a countable number of values; equals the pmf.
* continuous probability distribution: a probability distribution that takes an uncountable set of values; equals the pdf.
* cumulative distribution function: function that evaluates the probability that a random variable equals a value less than or equal to some value.
* likelihood: is a function that computes the probability of some data, fixed, given parameters that are random.

# Exercise 2)

## a)
First we should define our parameters $$\alpha$$ and $$\beta$$.

```{r}
mu <- .2
var <- .01
alpha <- mu*((mu*(1-mu))/var - 1)
beta <- alpha*(1 - mu)/mu
```

Now, we define our range of points over which we'll calculate and plot our density.

```{r}
data <- seq(.0001, .99, length.out = 10^3)
beta_density <- dbeta(data, alpha, beta)
plot(data, beta_density)
```

## b)

```{r}
draws <- rbeta(10^3, alpha, beta)
hist(draws, col = "blue")
```

The histogram is very similar visually to the density.

## c)

```{r}
mu_sample <- mean(draws)
var_sample <- var(draws)
```

The sample mean is `r mu_sample` while the true mean is 0.2.
The sample variance is `r var_sample` while the true variance is 0.01.

## d)

```{r}
quantile(draws, probs = c(.025, .975))
```

# Exercise 3)

p(test = positive | lcancer) = .98
p(test = positive | not_lcancer) = .04
p(test = negative | not_lcancer) = .96
p(lcancer) = .001

We need to find what is the probability that the subject has lung cancer, if he tested positive.
We want p(lcancer | test = positive).
By Bayes Rule, p(lcancer | test = positive) = p(lcancer, positive)/p(test = positive).

p(test = positive) = p(test = positive | lcancer)*p(lcancer) + p(test = positive | not_lcancer)*p(not_lcancer) = .98*.001 + .04*.999 = `r .98*.001 + .04*.999`.

Then, p(lcancer | test = positive) = p(lcancer, positive)/p(test = positive) = .98*.001/`r .98*.001 + .04*.999` = `r .98*.001/.98*.001 + .04*.999`.

If there is a high cost (in various dimensions) to the positive test, researchers should be aware that the low prevalence of the disease among the population, makes a positive test unreliable, a high false positive ratio.

# Exercise 4)

Create boxes.

```{r}
boxes <- matrix(c(2, 4, 1, 5, 1, 3), ncol = 2, dimnames = list(c("A", "B", "C"), c("RED", "WHITE")))
```

## a)

We want p(red).
p(red) = p(red | box = A)*p(A) + p(red | box = B)*p(B) + p(red | box = C)*p(C)

```{r}
p_red <- function(boxes) {
    p_a <- .4
    p_b <- .1
    p_c <- .5
    p_r <- (boxes["A", "RED"]/(boxes["A", "RED"] + boxes["A", "WHITE"]))*p_a + (boxes["B", "RED"]/(boxes["B", "RED"] + boxes["B", "WHITE"]))*p_b + (boxes["C", "RED"]/(boxes["C", "RED"] + boxes["C", "WHITE"]))*p_c
    p_r
}

p_red(boxes = boxes)
```

## b)

We want p(Box | red).
p(box = A | red) = p(box = A)*p(red | box = A)/p(red)

```{r}
p_box <- function(boxes) {
    p_a <- .4
    p_b <- .1
    p_c <- .5
    p_r <- p_red(boxes = boxes)
    p_ar <- (boxes["A", "RED"]/(boxes["A", "RED"] + boxes["A", "WHITE"]))*p_a/p_r
    p_br <- (boxes["B", "RED"]/(boxes["B", "RED"] + boxes["B", "WHITE"]))*p_b/p_r
    p_cr <- (boxes["C", "RED"]/(boxes["C", "RED"] + boxes["C", "WHITE"]))*p_c/p_r
    list(p_ar, p_br, p_cr)
}

p_box(boxes = boxes)
```

# Exercise 5

We want p(Elvis had identical twin | twin male).
p(Elvis had identical twin | twin male) = p(Elvis had identical twin)*p(twin male | Elvis had identical twin)/p(twin male)
= .0025*.5/(.0025*.5 + .0067*.5)
p(twin male) = p(identical twin)*p(male twin | identical twin) + p(fraternal twin)*p(male twin | fraternal twin) = .0025*.5 + .0067*.5

```{r}
p_identical_twin <- function(fraternal_prob, identical_prob) {
    f_p <- fraternal_prob
    i_p <- identical_prob
    p_male_i <- .5
    p_male_f <- .25
    p_id_tw <- i_p*p_male_i/(i_p*p_male_i + f_p*p_male_f)
    p_id_tw
}

p_identical_twin(fraternal_prob = 1/150, identical_prob = 1/400)
```