---
title: "Assignment 7"
author: "Anonymous"
date: "30/04/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(aaltobda)
library(rstan)
data("drowning")
```

# Linear model: drowning data with Stan

## 1.

```{stan output.var=""}
data {
  int<lower=0> N;
  vector[N] x;
  vector[N] y;
  real xpred;
}
parameters {
  real alpha;
  real beta;
  real<lower=0> sigma; //here I changed the upper limit = 0 to lower limit = 0.
}
transformed parameters {
  vector[N] mu = alpha + beta*x;
}
model {
  y ~ normal(mu, sigma); //there was a ; missing
}
generated quantities {
  real ypred = normal_rng(alpha + beta*xpred, sigma); //changed the pred to be based on xpred value.
}
```

The errors are reported in comments on the code.

## 2.

```{stan output.var=""}
data {
  int<lower=0> N; //data points
  vector[N] x; //obs year
  vector[N] y; //obs n of drowned
  real xpred; //prediction year
  real pmualpha; //prior mean alpha
  real pmubeta; //prior mean beta
  real psalpha; //prior std alpha
  real psbeta; //prior std beta
}
parameters {
  real alpha;
  real beta;
  real<lower=0> sigma;
}
transformed parameters {
  vector[N] mu = alpha + beta*x;
}
model {
  beta ~ normal(pmubeta, psbeta);
  y ~ normal(mu, sigma);
}
generated quantities {
  real ypred = normal_rng(alpha + beta*xpred, sigma);
}
```

For beta to lie in the range [-69, 69] with probability of 99%, the std has to be 26.8. I found that value by trial and error, based on a formula to output the interval.

## 3.

The code is in item 2 above. To implement the beta prior I declared the proper variables in data, line 6 through 9, and then in the model part I added beta following a distribution based on the prior values. I didn't write the values, pmubeta = 0 and psbeta = 26.8, because that would come in the RStan call.

## 4.

The code is the same as before, with only change in the first line of the model.

```{stan output.var=""}
model {
  alpha ~ normal(pmualpha, psalpha);
  beta ~ normal(pmubeta, psbeta);
  y ~ normal(mu, sigma);
}
```

As before, I didn't write in the code directly, because the values I'd select I'd put them in a RStan call.
My suggestion for alpha prior is that we model it following a normal density, with the approximate historical mean, 138, and standard deviation of 25; so pmualpha = 138 and psalpha = 25.
I selected that standard deviation value based on the visualization of the plot in Figure 1 provided in the assignment, so it would cover the entire range of intercepts.

# Hierarchical model: factory data with Stan

```{r}
data("factory")
sm <- stan_model(file = "~/Downloads/bda/bad_separate.stan")
stan_data <- list(y = factory, N = nrow(factory), J = ncol(factory))
s_model <- sampling(sm, data = stan_data)
```

